---
title: "Stat474W/574 Activity 2"
author: "Dr. Islam"
date: "Winter, Jan 20, 2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

# Objectives

- Review definitions of population and sample variances
- Compute population variance
- Compute sample variance
- Compute pooled sample variance for two populations with common variance
- Make graphical diaplays of data via boxplot and histogram

# 1. Useful definitions


**Definition 1.1** 

Given a finite population ${X_1,X_2,…,X_N}$, 

- the population mean, denoted by $\mu$, is defined by

$$\mu=\frac{\sum_{i=1}^NX_i}{N}$$

- the  population variance, denoted by $\sigma^2$ is defined by  

$$\sigma^2=\frac{1}{N}\sum_{i=1}^N(X_i-\mu_x)^2$$

**Definition 1.2**

Given a sample ${x_1,x_2,…,x_n}$ from a population with  mean $\mu$ and variance $\sigma^2$, 

- the sample mean, denoted by $\bar x$, is defined by

$$\bar x =\frac{\sum_{i=1}^nx_i}{n}$$

- the sample variance, denoted by $s^2$ is defined by  

$$s^2=\frac{1}{n-1}\sum_{i=1}^n(x_i-\bar x)^2$$

**Definition 1.3**

- Given two samples ${x_1,x_2,…,x_{n_1}}$ and ${y_1,y_2,…,y_{n_2}}$ from two populations with unknown but common variance $\sigma^2$, an estimate of the common variance $\sigma^2$ is given by

$s_{pooled}^2=\frac{\sum_{i=1}^{n_1}(x_i-\bar x)^2+\sum_{i=1}^{n_2}(y_i-\bar y)^2}{n_1+n_2-2}=\frac{(n_1-1)s_1^2+(n_2-1)s_2^2}{n_1+n_2-2}$

where $$s_1^2=\frac{1}{n_1-1}\sum_{i=1}^{n_1}(x_i-\bar x)^2$$
$$s_2^2=\frac{1}{n_2-1}\sum_{i=1}^{n_2}(y_i-\bar y)^2$$
$s^2_{pooled}$ is called the pooled sample variance of the unknown but common variance $\sigma^2$ of two populations.

# 2 Functions to be used

- sum()
- mean()
- var()
- sd()
- length()
- par()
- hist()
- boxplot()

# Example 1

The data below refers to a population of circumference of 7 orange trees, measured in mm, at breast height, a standard measurement in forestry.

```
209  58 203 145 201  51  75
```
 (a) Assign this population data values to a vector $X$
 (b) Find the population mean of circumference of trees ($\mu$).
 (c) Find the population variance of circumference of trees ($\sigma^2$).
 
Solution

There is no function in R to compute population variance $\sigma^2$. We can easily compute it via self-coding using known functions:
 
```{r}
#(a)
X<-scan(text="209  58 203 145 201  51  75")
N=7
#(b)
mu=mean(X)
#(c)
pop.var<-sum((X-mu)^2)/7
```

# Example 2

Below is sample of creative writing score of students of a writing class, after receiving intrinsic motivation:

```
19.3 21.3 22.6 20.3 17.5 22.2 26.7 29.7 20.5 20.6 22.1 18.2 12.0 19.1 23.1

```
(a) Compute sample variance by self-coding in R.
(b) The var() function in R computes sample variance. Compute sample variance using var() function.
(c) verify that the results in (a) and (b) agree.


```{r}
x<-scan(text="19.3 21.3 22.6 20.3 17.5 22.2 26.7 29.7 20.5 20.6 22.1 18.2 12.0 19.1 23.1")
n=length(x)
xbar<-mean(x)
#(a)
samp.var<-sum((x-xbar)^2)/(n-1)
# the function var() also compute sample variance
#(b)
samp.var2=var(x)
#(c)
c(samp.var, samp.var2)
```

# Example 3

Given two samples x and y below, x from intrinsic score and y from extrinsic score, as follows:

```
x<-scan(text="16.6 26.7 17.5 24.0 22.2 29.7 19.3 19.1 12.9 12.0 19.8 12.0")
y<-scan(text="17.2 15.0  6.1 11.8 19.2 16.8 14.8 12.3 12.0 24.0 20.7 17.2 10.9 22.1  5.0")
```

(a) Compute $s_1^2$, variance of the sample x.
(b) Compute $s_2^2$, variance of the sample y.
(c) Assume that the two populations of intrinsic and extrinsic scores have unknown but equal variances $\sigma^2$. Estimate pooled sample variance $s_{pooled}^2$ of the common variance $\sigma^2$.

```{r}
x<-scan(text="16.6 26.7 17.5 24.0 22.2 29.7 19.3 19.1 12.9 12.0 19.8 12.0")
y<-scan(text="17.2 15.0  6.1 11.8 19.2 16.8 14.8 12.3 12.0 24.0 20.7 17.2 10.9 22.1  5.0")
n1<-length(x)
n2<-length(y)
s1sq<-var(x)
s2sq<-var(y)
s2pooled<-((n1-1)*s1sq+(n2-1)*s2sq)/(n1+n2-2)
c(s1sq,s2sq,s2pooled)
par(mfrow=c(2,2))#row-wise fill
boxplot(x, main="Boxplot of x")
boxplot(y, main="Boxplot of y")
hist(x)
hist(y)
par(mfcol=c(2,2))#column-wise fill
boxplot(x, horizontal = T, main="Boxplot of x")
boxplot(y, horizontal = T, main="Boxplot of y")
hist(x)
hist(y)
```



