---
title: "Final574-LianZuo"
author: "LianZuo"
date: '`r Sys.Date()`'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment="")
```

# Problem 1

Test the randomness of each of the following sequences:

(i) 1 2 3 4 5 6 7 4 4 4 4 4 4  

(ii) T F T F T T F T T T F T T T

(iii)  D D D D D D D D N N N

(iv) 12 13 11 11  8 12 12  9  9  5  8  8

(v) -1  1 -1 -1  1  1 -1 -1  1 -1  1 -1
at significance level α=0.05

Solution:
```{r}
# (i)
x1 <- scan(text="1 2 3 4 5 6 7 4 4 4 4 4 4")
DescTools::RunsTest(x1)
# (ii)
x2 <- c("T","F", "T", "F", "T","T","F", "T","T","T","F","T","T","T","F")
DescTools::RunsTest(x2)
# (iii)
x3 <- c("D","D", "D", "D", "D","D","D", "D","N","N","N")
DescTools::RunsTest(x3)
# (iv)
x4 <- scan(text="12 13 11 11  8 12 12  9  9  5  8  8")
DescTools::RunsTest(x4)
# (v)
x5 <- scan(text="-1  1 -1 -1  1  1 -1 -1  1 -1  1 -1")
DescTools::RunsTest(x5)
```

# Problem 2

Given the following sequence of number

 9 8 9 3 9 9 6 4 9 6 3 8 8 5 8
(i) Determine the number of runs by comparing each number with the medians.

(ii) Determine the number of runs by comparing each with the previous number.

Solution:
```{r}
x <- scan(text=" 9 8 9 3 9 9 6 4 9 6 3 8 8 5 8")
# (i)
med=median(x)
run1=x>med
run1
run2=as.numeric(x>med)
run2
DescTools::RunsTest(x>med)
# (ii)
y <- diff(x)
length(y)
DescTools::RunsTest(y)
z <- ifelse(y>0,"+", "-")
length(z)
z
```

# Problem 3

The failure time T of a certain product follows an exponential distribution with mean of 3075 days.

(a) Find the MTTF of the product.

(b) Find the reliability of the product at t=2000 days.

(c) Find the probability that the item will survive to its MTTF.

Solution:
```{r}
mu=3075
lambda=1/mu
t=2000
#(a)
## MTTF=u
MTTF=mu
#(b) 
##R(t)=1-F(t)
Rt=exp(-lambda*t)
#(c)
RMTTF=exp(-lambda*mu)
```

# Problem 4

A production system consists of three 3 independent components C1,C2,C3 whose time to failures follow exponential distributions with rates λ1=2500,λ2=2650,λ3=2475 days.

Compute the probability that component C2 will fail first.

Solution:
```{r}
lambda1=2500
lambda2=2650
lambda3=2475
P2FF=lambda2/(lambda1+lambda2+lambda3)
```

# Problem 5

Time to failure of a diesel engine compressor cooling system follows an Weibull distribution with a shape parameter α=2.4 and a scale parameter β=1973. Compute the following quantities:

(a) Compute the reliability function and evaluate it at t=1820. 

(b) Compute the mean time to failure (MTTF)

(c) Compute the mean residual life at time t and evaluate it at t=MTTF.

(d) Write the expression for the failure rate function and evaluate it at t=MTTF

Solution:
```{r}
t=1820
alpha=2.4
beta=1973
#(a)
RF=exp(-(t/beta)^alpha)
pweibull(1820, shape=alpha, scale=beta, lower.tail=F)
#(b)
MTTF=beta*gamma((1/alpha)+1)
#(c)
MRL.MTTF=(beta/alpha)*(MTTF/beta)^(alpha-1)
#(d)
z.MTTF=(alpha/beta)*(MTTF/beta)^(alpha-1)

```

# Problem 6

The data vector below is assumed to come from an Weibull(α,β) distribution with α and β unknown.

x=scan(text="1.51 1.41 1.75 1.32 1.30 0.84 1.17 0.65 1.26 1.14 1.74 1.00 1.29 0.89 1.81 0.67 0.63 1.20 1.92 1.33 0.64 1.59 1.09 1.49 1.72 1.39 0.97 1.67 0.73 1.58 0.95 1.12 1.01 1.28 0.96 1.10 1.47 0.83 0.66 1.08 0.52 0.50 0.87 1.43 0.38")

(a) Estimate α and β using any of the R function discussed in the class. 

(b) Use the estimated parameters to test if the data actually come from the hypothesized Weibull distribution.

Solution:

```{r}
x <- scan(text="1.51 1.41 1.75 1.32 1.30 0.84 1.17 0.65 1.26 1.14 1.74 1.00 1.29 0.89 1.81 0.67 0.63 1.20 1.92 1.33 0.64 1.59 1.09 1.49 1.72 1.39 0.97 1.67 0.73 1.58 0.95 1.12 1.01 1.28 0.96 1.10 1.47 0.83 0.66 1.08 0.52 0.50 0.87 1.43 0.38")
fit=MASS::fitdistr(x, "weibull",lower=0)
fit
ks.test(x,"pweibull", scale=fit$estimate[2], shape=fit$estimate[1])
``` 

# Problem 7

Given the following ordered pair (x,y):


x: 14.8 13.3 14.9 14.5 13.9 15.7 12.7 16.5 16.6 15.3 14.5 16.6 12.8 15.8 12.8
y: 15.2 12.3 14.5 17.6 16.0 14.5 14.1 15.1 10.9 15.7 13.5 11.4 12.4 16.2 12.8

(a) Test H0:μx−μy=0 against the two sided alternative Ha:μx−μy≠0 using any appropriate test.

(b) Perform a permutation test for the above test using 5000 permutation samples.

(c) What is your conclusion about the test? Use significance level α=0.05

Solution:

```{r}
# (a)
x <- scan(text="14.8 13.3 14.9 14.5 13.9 15.7 12.7 16.5 16.6 15.3 14.5 16.6 12.8 15.8 12.8")
y <- scan(text="15.2 12.3 14.5 17.6 16.0 14.5 14.1 15.1 10.9 15.7 13.5 11.4 12.4 16.2 12.8")

t.test(x,y, paired=T, alpha=0.05)

# (b)
set.seed(2022)
difference=x-y
n=length(difference)
obs.diff=mean(difference)
perm.mean.diff=c()
for (i in 1:5000){
  perm.mean.diff[i]=mean(difference*sample(c(-1,1),n, rep=T))
}
pvalue=mean(abs(perm.mean.diff)>abs(obs.diff))
pvalue

```